=enter title here=

### 2021-08-08 paper/agujournaltemplate.tex
Figures for MODWT

In this study, we use wavelet methods to analyze GPS and seismic recordings of slow slip events in Cascadia. Our objective is to verify that there is a good correlation between slow slip events detected with only GPS data, and slow slip events detected with only seismic data. We thus want to demonstrate that the wavelet-based detection method can be applied to detect possible smaller (magnitude 5) slow slip events that may be currently undetected with standard methods.

@@

@@

The wavelet methods for time series analysis are explained in a more detailed way in Percival \& Walden (2000 ~\cite{PER_2000}). \\

\textbf{Discrete Wavelet Transform}

The Discrete Wavelet Transform (DWT) is an orthonormal transform that transforms a time series $X_t \left( t = 0, ... , N - 1 \right)$ into a vector of wavelet coefficients $W_i \left( i = 0 , ... , N - 1 \right)$. If we denote $J$ the level of the wavelet decomposition, and we have $N = n* 2^J$, where n is some integer higher or equal to 1, the vector of wavelet coefficients can be decomposed into $J$ wavelet vectors $W_j$ of lengths $\frac{N}{2}$, $\frac{N}{4}$, ... , $\frac{N}{2^J}$, and one scaling vector $V_J$ of length $\frac{N}{2^J}$. \\

Each wavelet vector $W_j$ is associated with changes on scale $\tau_j = dt 2^{j - 1}$, where $dt$ is the time step of the time series, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack \frac{1}{dt 2^{j + 1}} ; \frac{1}{dt 2^j} \rbrack$. The scaling vector $V_J$ is associated with averages in scale $\lambda_J = dt 2^J$, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack 0 ; \frac{1}{dt 2^{j + 1}} \rbrack$. \\

We can also define for $j = 1 , ... , J$ the $j$th wavelet detail $D_j$, which is a vector of length $N$, and is associated to scale $\tau_j = dt 2^{j - 1}$. Similarly, we can define for $j = 1 , ... , J$ the $j$th wavelet smooth $S_j$, which is a vector of length $N$, and is associated to scales $\tau_{j + 1} = dt 2^{j + 1}$ and higher. Together, the details and the smooths define the multiresolution analysis (MRA) of $X$:

\begin{equation}
X = \sum_{j = 1}^{J} D_j + S_J
\end{equation}

One main advantage of the DWT is that it is an orthonormal transform, and thus we can write the analysis of variance (ANOVA):

\begin{equation}
\left\Vert X \right\Vert ^2 = \left\Vert W \right\Vert ^2 = \sum_{j = 1}^{J} \left\Vert W_j \right\Vert ^2 + \left\Vert V_J \right\Vert ^2 = \sum_{j = 1}^{J} \left\Vert D_j \right\Vert ^2 + \left\Vert S_J \right\Vert ^2
\end{equation}

Moreover, the DWT can be computed using $O \left( N \right)$ multiplications. \\

However, the DWT present several disadvantages:

\begin{itemize}
	\item The length of the time series must be a multiple of $2^J$ where $J$ is the level of the DWT decomposition.
	\item The time step of the wavelet vector $W_j$ is $dt 2^j$, which may not correspond to the time when some interesting phenomenon is visible on the original time series.
	\item When we circularly shift the time series, the corresponding wavelet coefficients, details and smooths are not a circularly shifted version of the wavelet coefficients, details and smooths of the original time series. Thus, the values of the wavelet coefficients, details and smooths are strongly dependent on the time when we start experimentally gathering the data.
	\item When we filter the time series to obtain the details and smooths, we introduce a phase shift, which makes difficult to line up meaningfully the features of the MRA with the original time series.
\end{itemize}

\textbf{Maximum Overlap Discrete Wavelet Transform}

To get rid of these problems, we introduce the Maximum Overlap Discrete Wavelet Transform (MODWT). The MODWT transforms the time series $X_t \left( t = 0, ... , N - 1 \right)$ into J wavelet vectors $\widetilde{W}_j \left( j = 1 , ... , J \right)$ of length $N$ and a scaling vector $\widetilde{V}_J$ of length $N$. As is the case for the DWT, each wavelet vector $\widetilde{W}_j$ is associated with changes on scale $\tau_j = dt 2^{j - 1}$, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack \frac{1}{dt 2^{j + 1}} ; \frac{1}{dt 2^j} \rbrack$. The scaling vector $\widetilde{V}_J$ is associated with averages in scale $\lambda_J = dt 2^J$, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack 0 ; \frac{1}{dt 2^{j + 1}} \rbrack$. \\

As is the case for the DWT, we can write the MRA:

\begin{equation}
X = \sum_{j = 1}^{J} \widetilde{D}_j + \widetilde{S}_J
\end{equation}

and the ANOVA:

\begin{equation}
\left\Vert X \right\Vert ^2 = \sum_{j = 1}^{J} \left\Vert \widetilde{W}_j \right\Vert ^2 + \left\Vert \widetilde{V}_J \right\Vert ^2
\end{equation}

Now, we have the following properties:

\begin{itemize}
	\item The MODWT of a time series can be defined for any length $N$.
	\item The time step of the wavelet vectors $\widetilde{W}_j$ and the scaling vector $\widetilde{V}_J$ is equal to the time step of the original time series.
	\item When we circularly shift the time series, the corresponding wavelet vectors, scaling vector, details and smooths are shifted by the same amount.
	\item The details and smooths are associated with a zero phase filter, making it easy to line up meaningfully the features of the MRA with the original time series.
\end{itemize}

However, the MODWT has some disadvantages over the DWT:

\begin{itemize}
	\item The MODWT can only be computed using $O \left( N \log_2 N \right)$ multiplications.
	\item We can no longer write the ANOVA for the details and smooths:

\begin{equation}
\left\Vert X \right\Vert ^2 \neq \sum_{j = 1}^{J} \left\Vert D_j \right\Vert ^2 + \left\Vert S_J \right\Vert ^2 \text{ and } \left\Vert \widetilde{W}_j \right\Vert ^2 \neq \left\Vert \widetilde{D}_j \right\Vert ^2
\end{equation}

\end{itemize}

To illustrate the wavelet transform method, we first apply the MODWT to synthetics data. As slow slip events occur in Cascadia on a regular basis, every twelve to eighteen months, we create a synthetic signal of period $T = 500$ days. To reproduce the ground displacement observed on the longitudinal component of GPS stations in Cascadia, we divide each period into two parts: In the first part of duration $T - N$, the displacement is linearly increasing and corresponds to the secular plate motion in the eastern direction; in the second part of duration $N$, the displacement is linearly decreasing and corresponds to a slow slip event on a reverse fault at depth triggering a ground displacement in the western direction. To see the effect of the magnitude of the slow slip event, we use different values for $N = 2, 5, 10, 20$ days. Figure X shows the synthetics, the details of the wavelet decomposition for levels 1 to 8, and the smooth for the four durations of a slow slip event. \\

%\begin{figure}
%\noindent\includegraphics[width=\textwidth, trim={5cm 6cm 5cm 9cm},clip]{figures/500_DS.pdf}
%\caption{Details and smooth of the wavelet decomposition of a synthetics signal with period 500 days and duration of the slow slip event equal to 2 days (left), 5 days, 10 days, and 20 days (right).}
%\label{pngfiguresample}
%\end{figure}

@@

>>>>

In this study, we use wavelet methods to analyze GPS and seismic recordings of slow slip events in Cascadia. Our objective is to verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data. We thus want to demonstrate that the wavelet-based detection method can be applied to detect possible smaller (magnitude 5) slow slip events that may be currently undetected with standard methods. \\

@@

We focused our study on northwest Washington State. For the GNSS data, we used the GPS time series provided by the Pacific Northwest Geodetic Array, Central Washington University. These are network solutions in ITRF2008 with phase ambiguities resolved. Solutions are computed with JPL/NASA orbits and satellite clocks. North, East, and Vertical directions are available. However, as the direction of the secular plate motion is close to the East direction, we only used the East direction of the GPS time series for the data analysis, as it has the best signal-to-noise ratio. The wavelet method works best with data with zero mean, and no sharp discontinuities, so we use the cleaned dataset, that is GPS times eries with linear trends, steps due to earthquakes or hardware upgrades, and annual and semi-annual sinusoids signals simultaneously estimated and removed following ~\citet{SZE_2004}. For the seismic data, we used the tremor catalog from the Pacific Northwest Seismic Network (PNSN) ~\citep{WEC_2010}. Tremor were detected and located using waveform envelope correlation and clustering and a centroid location is available for every given five-minute time window when tremor was detected. As the catalog starts in August 2009, we only looked at GPS data recorded in 2009 or later. \\

@@

\subsection{The Maximal Overlap Discrete Wavelet Transform}

The Discrete Wavelet Transform (DWT) is an orthonormal transform that transforms a time series $X_t \left( t = 0, \cdots , N - 1 \right)$ into a vector of wavelet coefficients $W_i \left( i = 0 , \cdots , N - 1 \right)$. If we denote $J$ the level of the wavelet decomposition, and we have $N = n * 2^J$, where $n$ is some integer higher or equal to 1, the vector of wavelet coefficients can be decomposed into $J$ wavelet vectors $W_j$ of lengths $\frac{N}{2}$, $\frac{N}{4}$, ... , $\frac{N}{2^J}$, and one scaling vector $V_J$ of length $\frac{N}{2^J}$. Each wavelet vector $W_j$ is associated with changes on scale $\tau_j = dt 2^{j - 1}$, where $dt$ is the time step of the time series, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack \frac{1}{dt 2^{j + 1}} ; \frac{1}{dt 2^j} \rbrack$. The scaling vector $V_J$ is associated with averages in scale $\lambda_J = dt 2^J$, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack 0 ; \frac{1}{dt 2^{j + 1}} \rbrack$. We can also define for $j = 1 , \cdots , J$ the $j$th wavelet detail $D_j$, which is a vector of length $N$, and is associated to scale $\tau_j = dt 2^{j - 1}$. Similarly, we can define for $j = 1 , \cdots , J$ the $j$th wavelet smooth $S_j$, which is a vector of length $N$, and is associated to scales $\tau_{j + 1} = dt 2^{j + 1}$ and higher. Together, the details and the smooths define the multiresolution analysis (MRA) of $X$:

\begin{linenomath*}
\begin{equation}
X = \sum_{j = 1}^{J} D_j + S_J
\end{equation}
\end{linenomath*}

The DWT present several disadvantages. First, the length of the time series must be a multiple of $2^J$ where $J$ is the level of the DWT decomposition. Second, the time step of the wavelet vector $W_j$ is $dt 2^j$, which may not correspond to the time when some interesting phenomenon is visible on the original time series. Third, when we circularly shift the time series, the corresponding wavelet coefficients, details and smooths are not a circularly shifted version of the wavelet coefficients, details and smooths of the original time series. Thus, the values of the wavelet coefficients, details and smooths are strongly dependent on the time when we start experimentally gathering the data. Finally, when we filter the time series to obtain the details and smooths, we introduce a phase shift, which makes difficult to line up meaningfully the features of the MRA with the original time series. \\

This is why we use instead the Maximal Overlap Discrete Wavelet Transform (MODWT). The MODWT transforms the time series $X_t \left( t = 0, ... , N - 1 \right)$ into J wavelet vectors $\widetilde{W}_j \left( j = 1 ,  \cdots , J \right)$ of length $N$ and a scaling vector $\widetilde{V}_J$ of length $N$. As is the case for the DWT, each wavelet vector $\widetilde{W}_j$ is associated with changes on scale $\tau_j = dt 2^{j - 1}$, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack \frac{1}{dt 2^{j + 1}} ; \frac{1}{dt 2^j} \rbrack$. The scaling vector $\widetilde{V}_J$ is associated with averages in scale $\lambda_J = dt 2^J$, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack 0 ; \frac{1}{dt 2^{j + 1}} \rbrack$. As is the case for the DWT, we can write the MRA:

\begin{linenomath*}
\begin{equation}
X = \sum_{j = 1}^{J} \widetilde{D}_j + \widetilde{S}_J
\end{equation}
\end{linenomath*}

The MODWT of a time series can be defined for any length $N$. The time step of the wavelet vectors $\widetilde{W}_j$ and the scaling vector $\widetilde{V}_J$ is equal to the time step of the original time series. When we circularly shift the time series, the corresponding wavelet vectors, scaling vector, details and smooths are shifted by the same amount. The details and smooths are associated with a zero phase filter, making it easy to line up meaningfully the features of the MRA with the original time series. The wavelet methods for time series analysis are explained in a more detailed way in ~\citet{PER_2000}). \\

\subsection{Application to synthetic data}

To illustrate the wavelet transform method, we first apply the MODWT to synthetics data. As slow slip events occur in Cascadia on a regular basis, every twelve to eighteen months, we create a synthetic signal of period $T = 500$ days. To reproduce the ground displacement observed on the longitudinal component of GPS stations in Cascadia, we divide each period into two parts: In the first part of duration $T - N$, the displacement is linearly increasing and corresponds to the secular plate motion in the eastern direction; in the second part of duration $N$, the displacement is linearly decreasing and corresponds to a slow slip event on a reverse fault at depth triggering a ground displacement in the western direction. To see the effect of the magnitude of the slow slip event, we use different values for $N = 2, 5, 10, 20$ days. Figure 1 shows the synthetics, the details of the wavelet decomposition for levels 1 to 8, and the smooth for the four durations of a slow slip event. \\

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/500_DS.eps}
\caption{Details and smooth of the wavelet decomposition of a synthetics signal with period 500 days and duration of the slow slip event equal to 2 days (left), 5 days, 10 days, and 20 days (right).}
\label{pngfiguresample}
\end{figure}

@@

\subsection{MODWT of GPS and tremor data}
